Notes on the Godunov Step in {\tt MAESTRO} and {\tt VARDEN}

### 2008-04-21 docs/Godunov_notes/paper.tex
updated to account for latest cvs checkins.  Document is now consistent with code

These are working notes for the Godunov step in {\tt MAESTRO} and {\tt VARDEN}.
\section{Differences Between These Notes and the Code}
\begin{enumerate}
\item Concerning both equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}), in the code we are omitting the terms in equation (\ref{Perturbational Velocity Equation}) proportional to $w_0$ (by using $\wt$ rather than $w$ as the advection speed) and $(\Ubt\cdot\eb_r)$ (by omitting the term altogether).
\item Concerning (\ref{Force Term}), in the code we are not including the term proportional to $\partial\pi_0/\partial r$ in the edge-state prediction.  However, this term is present in the cell-update.
\end{enumerate}
\section{Questions}
\begin{enumerate}
\item We need to re-examine when to include the term proportional to $(\Ubt\cdot\eb_r)$ in equation (\ref{Perturbational Velocity Equation}) when we do a 1D extrapolation of the radial velocity to r-faces.  For example, equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}) include it, but equations (\ref{1D wt_L extrap}) and (\ref{1D wt_R extrap}) do not.
\item In equations (\ref{wt_L Extrapolation}) and (\ref{wt_R Extrapolation}), (\ref{wt_L x-edge}) and (\ref{wt_R x-edge}), (\ref{wt_L r-edge}) and (\ref{wt_R r-edge}), we could also compute the term proportional to $\partial w_0/\partial r$ by using the cell-centered $\wt$ rather than the average of the neighboring $\wt^{\trans}$.
\end{enumerate}
\section{Differences Between {\tt MAESTRO} and {\tt VARDEN}}
\begin{enumerate}
\item {\tt VARDEN} does not use the concept of $w_0$, so $\Ubt$ does not exist.
\item {\tt VARDEN} only does Cartesian problems.
\item {\tt VARDEN} does not use the concept of $\Ub^{\trans}$, but rather in {\tt advance\_premac}, the function {\tt velpred} computes $\Ub^{\mac,*}$ directly.
\item {\tt VARDEN} uses characteristic tracing in the Taylor series extrapolation for $\Ub^{\mac,*}$, whereas {\tt MAESTRO} does not.
\item {\tt VARDEN} uses an ``FPU'' type algorithm to get both $\Ub^{\edge}$ and $s^{\edge}$, whereas {\tt MAESTRO} uses the ``FPU'' algorithm for $s^{\edge}$, and a ``non-FPU'' algorithm for $\Ub^{\edge}$
\item {\tt VARDEN} treats $\rho$ in a conservative manner, whereas {\tt MAESTRO} treats $\rho'$ in a non-conservative manner.  Therefore, {\tt MAESTRO} accounts for this difference by modifying the forcing term in {\tt modify\_scal\_force}
\item {\tt MAESTRO} does not corner couple for 3D problems.
\end{enumerate}
\cleardoublepage
\section{Computing $\Ubt^{\trans}$ in {\tt MAESTRO}}
In {\tt advance\_premac}, we call the function {\tt mkutrans}, which computes the edge-centered transverse velocities, $\Ubt^{\trans}$.  These transverse velocities do not contain $w_0$, so immediately following the call to {\tt mkutrans}, we call {\tt addw0} to compute $\Ub^{\trans}$ from $\Ubt^{\trans}$.  We will only compute the normal component of velocity at each face.\\

@@

\wt_{L,\ib-\half\eb_r}^{1D} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r \wt_{\ib-\eb_r}^n, \label{1D wt_L extrap} \\
\wt_{R,\ib-\half\eb_r}^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n. \label{1D wt_R extrap}

>>>>

These are working notes for the Godunov step in {\tt MAESTRO} and {\tt VARDEN}.
\section{Differences Between These Notes and the Code}
\begin{enumerate}
\item Concerning equations (\ref{Radial utrans_L}), (\ref{Radial utrans_R}), (\ref{1D wt_L extrap}) and (\ref{1D wt_R extrap}), in the code we are omitting the term in equation (\ref{Perturbational Velocity Equation}) proportional to $(\Ubt\cdot\eb_r)$.
\item Concerning (\ref{Force Term}), in the code we are not including the term proportional to $\partial\pi_0/\partial r$ in the edge-state prediction.  However, this term is present in the cell-update.
\end{enumerate}
\section{Questions}
\begin{enumerate}
\item Concerning equations (\ref{Radial utrans_L}), (\ref{Radial utrans_R}), (\ref{1D wt_L extrap}) and (\ref{1D wt_R extrap}), do we really need to include the term proportional to $(\Ubt\cdot\eb_r)$?
\item Concerning (\ref{Force Term}), do we really need to include the term proportional to $\partial\pi_0/\partial r$ in the edge-state prediction?  If so, do we use the $\pi_0$ or $w_0$ formulation (refer to Paper III equation [38])?
\item Concerning equations (\ref{wt_L Extrapolation}) and (\ref{wt_R Extrapolation}), (\ref{wt_L x-edge}) and (\ref{wt_R x-edge}), (\ref{wt_L r-edge}) and (\ref{wt_R r-edge}), we could also compute the term proportional to $\partial w_0/\partial r$ by using the cell-centered $\wt$ rather than the average of the neighboring $\wt^{\trans}$.
\end{enumerate}
\section{Differences Between {\tt MAESTRO} and {\tt VARDEN}}
\begin{enumerate}
\item {\tt VARDEN} does not use the concept of $w_0$, so $\Ubt$ does not exist.
\item {\tt VARDEN} only does Cartesian problems.
\item {\tt VARDEN} does not use the concept of $\Ub^{\trans}$, but rather in {\tt advance\_premac}, the function {\tt velpred} computes $\Ub^{\mac,*}$ directly.
\item {\tt VARDEN} uses characteristic tracing in the Taylor series extrapolation for $\Ub^{\mac,*}$, whereas {\tt MAESTRO} does not.
\item {\tt VARDEN} uses an ``FPU'' type algorithm to get both $\Ub^{\edge}$ and $s^{\edge}$, whereas {\tt MAESTRO} uses the ``FPU'' algorithm for $s^{\edge}$, and a ``non-FPU'' algorithm for $\Ub^{\edge}$
\item {\tt VARDEN} treats $\rho$ in a conservative manner, whereas {\tt MAESTRO} treats $\rho'$ in a non-conservative manner.  Therefore, {\tt MAESTRO} accounts for this difference by modifying the forcing term in {\tt modify\_scal\_force}
\item {\tt MAESTRO} does not corner couple for 3D problems.
\end{enumerate}
\cleardoublepage
\section{Computing $\Ubt^{\trans}$ in {\tt MAESTRO}}
In {\tt advance\_premac}, we call the function {\tt mkutrans}, which computes the edge-centered transverse velocities, $\Ubt^{\trans}$.  These transverse velocities do not contain $w_0$, so immediately following the call to {\tt mkutrans}, we call {\tt addw0} to compute $\Ub^{\trans}$ from $\Ubt^{\trans}$.  We will only compute the normal component of velocity at each face.\\

@@

\wt_{L,\ib-\half\eb_r}^{1D} &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{2h}w_{\ib-\eb_r}^n\right)\Delta_r \wt_{\ib-\eb_r}^n - \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_r}, \label{1D wt_L extrap} \\
\wt_{R,\ib-\half\eb_r}^{1D} &=& \wt_{\ib} - \left(\half + \frac{\dt}{2h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n - \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib}. \label{1D wt_R extrap}
